Growing up with Nell: A Narrative Interface for Literacy

### 2012-03-08 paper.tex
Intro story.

\begin{abstract}
Project Nell%
\footnote{We were inspired by the ``Young Lady's Illustrated
Primer'' in the Neal Stephenson novel \textit{The Diamond Age}.
Nell takes its name from that novel's protagonist, but could also be read as an
acronym for ``Narrative Environment for Learning Learning'' (a nod to
Seymour Papert).}
is tablet software to teach reading and writing to
children far from educational infrastructure.  Nell is a modular
narrative system with ``a low floor and no ceilings'': the design
scales from teaching letter shapes to programming the system itself.

@@

\section{A Lesson from Nell}
% tell a nell story up front

\section{Key Ideas}
% olpc context here?
\cite{chang:tinkrbook, flores:uruguay, idb:peru}
%context: english literacy, far from teachers & schools
\subsection{Narrative}
% excerpt of story model
\subsection{Personal}
% customization, renaming characters
\subsection{Direct interaction}
% handwriting as UI?
% motor skills
% photo of XO-3?
\subsection{Growable (``no ceiling'')}
% modular -- the story expands.
% excerpt of TurtleScript
% text read aloud initially

>>>>

\begin{abstract}
Project Nell%

is tablet software to teach reading and writing to
children far from educational infrastructure.  Nell is a modular
narrative system with ``a low floor and no ceilings'': the design
scales from teaching letter shapes to programming the system itself.

@@

\section{A Lesson from Nell}

% tell a nell story up front
Miles from the nearest school, a young girl in Ethiopia turns on a
tablet computer for the first time.  The solar-powered machine begins
talking to her immediately: \textit{``Hello!  Would you like to hear a story?''}

She listens intently to a story about a princess named Nell.  Later, when
the girl has learned a little more, she will give the princess her own
name, ``Rahel,'' and change all the colors, but for now the green
book shows her pictures and asks her to trace shapes on the screen.

\textit{``R is for Raven.  Can you trace the R?''}

As she traces the R, it comes to life and gallops across the screen.
\textit{``Run starts with R.  Roger the R Runs across the Red Rug.  He has a
dog named Rover.''}  Rover barks: \textit{``Ruff! Ruff!''}  The Princess asks,
\textit{``Can you find something Red?''} and Rahel uses the camera to
photograph her red boots. \textit{``Very good!  You are a smart girl!''}

As Rahel grows, the book asks her to trace not just letters, but whole
words.   The books responses are written on the screen as it writes
them, and eventually she doesn't need to leave the sound on all the
time to read what the Princess is saying.  Soon Rahel can write
complete sentences in her special book, and the Princess will respond
sometimes.  New stories teach her about music (by playing certain
tunes she unlocks a dungeon door) and programming with blocks (a
not-very-bright turtle needs her to tell him how to draw different
shapes).  Rahel writes her own stories, which she shares with her
far-away friends.

An older Rahel learns that the block language she used to talk with
the turtle is actually used to write all the software running inside
her special book.  The stories Rahel writes can now include new
functionality that make her book even more exciting for the kids
just now opening their green meta books.

\section{Key Ideas}
The interaction design of the Nell system described above is inspired
by the ``Young Lady's Illustrated Primer'' in the Neal Stephenson
novel \textit{The Diamond Age}, from whose protagonist Nell takes its
name.\footnote{In a nod to Seymour Papert, Nell can also be read as an
acronym for ``Narrative Environment for Learning Learning.''}
Nell's design embodies four key ideas: it is a \textbf{Narrative}
interface which is \textbf{Personalized} for the child and uses
\textbf{Direct Interaction}.  Nell is \textbf{Growable}: it has a
``low floor and no ceiling'' to allow a child to master more
and more of the system as she grows.

% olpc context here?
\cite{chang:tinkrbook, flores:uruguay, idb:peru}
%context: english literacy, far from teachers & schools
\subsection{Narrative}
% excerpt of story model
\subsection{Personal}
% customization, renaming characters
\subsection{Direct interaction}
% handwriting as UI?
% motor skills
% photo of XO-3?
\subsection{Growable (``no ceiling'')}
% modular -- the story expands.
% excerpt of TurtleScript
% text read aloud initially
